---
title: "Cluster analysis in segmenting ICT Skills among youths and adults"
author: "Molo Muli"
date: '2022-07-07'
output: 
  tufte::tufte_html: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r R-Script file, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
source("Scripts/ICT.Skills.Cluster.R", local = knitr::knit_global())
```


# Context

The data context falls under United Nations SDG 4 which is to ensure inclusive and equitable education and promote lifelong learning opportunities for all. Observations in the data entails percentage values of youths/adults who used 9 selected ICT skills. An individual is considered to have ICT skills if s/he used at least one of the 9 ICT skills in the past 3 months. The data set was collected using [Multiple Indicator Cluster Survey (MICS)](https://mics.unicef.org/) and it's hosted in [UNICEF's open data portal website](https://data.unicef.org/resources/dataset/education-data/).

*Disclaimer: All reasonable precautions have been taken to verify the information in this database. In no event shall UNICEF be liable for damages arising from its use or interpretation*

# Aim

To evaluate methods of clustering using a data set that contains both numerical and categorical observations

# Goals/Objectives

-   Discover similar or dissimilar ICT cluster skills across men and women in the various countries present in the data
-   R-implementation of different clustering techniques using a data set with mixed data types. A data set with mixed data types is a set of data containing both continuous and discrete observations.

# Data
The data set contains 15 columns and 285 observations. 73% is numerical and 27% discrete. Sampled countries are 35 across 8 UNICEF Regions. Below are the columns present in the data 

# Descriptive Analytics


From the data table above, the ICT skill with the highest variation is how to write a computer program using any programming language. Personally this is a very specific skill set as compared to other general ICT skills.


```{r Descriptive Statistics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Figure: Here is a really important caption."}
FormattableND
```


# Cluster analysis
Clustering is an unsupervised machine learning algorithm that finds clusters such that observations present in a cluster are similar to each other while clusters formed from these observations are dissimilar from each other.





Below is the

# Clustering

K-Modes Gowers Distance K-prototypes Jaccard Distance

# 

K-modes, unlike K-means uses dissimilarities, i.e quantification of total mismatches between two objects, the smaller the number the

# References

-   <https://link.springer.com/article/10.1023/A:1009769707641>
-   <https://amva4newphysics.wordpress.com/2016/10/26/into-the-world-of-clustering-algorithms-k-means-k-modes-and-k-prototypes/comment-page-1/>
