---
title: "Cluster analysis of mixed data types: Case study of segmenting ICT Skills among youths and adults"
author: "Molo Muli"
date: "`r Sys.Date()`"
output: 
  tufte::tufte_html: default
  code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r R-Script file, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
source("Scripts/ICT.Skills.Cluster.R", local = knitr::knit_global())
```


# Context

The data context falls under United Nations SDG 4 which is to ensure inclusive and equitable education and promote lifelong learning opportunities for all. Observations in the data entails percentage values of youths/adults who used 9 selected ICT skills. An individual is considered to have ICT skills if s/he used at least one of the 9 ICT skills in the past 3 months. The data set was collected using [Multiple Indicator Cluster Survey (MICS)](https://mics.unicef.org/) and it's hosted in [UNICEF's open data portal website](https://data.unicef.org/resources/dataset/education-data/).

*Disclaimer: All reasonable precautions have been taken to verify the information in this database. In no event shall UNICEF be liable for damages arising from its use or interpretation*

# Goals/Objectives

-   Cluster similar or dissimilar ICT skills across men and women in the various countries present in the data
-   Successfully use R to implement clustering on a data set with mixed data types. A data set with mixed data types is a set of data containing both continuous and discrete observations.

# Data
The data set contains 15 columns and 285 observations. 73% is numerical and 27% discrete. Sampled countries are 35 across 8 UNICEF Regions.

# Descriptive Analytics

```{r ND, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.margin=T}
NumericalDistribution
```


From the data table above, the ICT skill with the highest variation is how to write a computer program using any programming language. Personally this is a very specific skill set as compared to other general ICT skills.
From the histograms, all numeric variables have a right-tailed distribution. A probable reason is because the uptake of these courses was high initially, but over the period of three months, it slowed down. This is called the startup effect[^1].

```{r Descriptive Statistics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
FormattableND
```

<!-- # Exploratory Data Analysis (EDA) -->
<!-- ## Correlation Analysis -->
<!-- Correlation analysis is used to quantify the association of two continuous variables. Output of correlation analysis is a correlation coefficient which is a value that ranges from -1 to 1. This coefficient denotes the strength of the association of the variables. Variables which have a coefficient leaning towards -1 have a negative correlation whereas variables that lean towards a coefficient value of 1 have a positive correlation. The higher the coefficient on either sides, the stronger the correlation between the variables. A correlation coefficient of 0 means that the variables are independent of each other. -->

<!-- ## R-Implementation of correlation -->
<!-- Since correlation works with numeric values, subset the data frame to remain with only continuous values -->

# Cluster analysis
Clustering is an unsupervised machine learning algorithm that finds clusters such that observations present in a cluster are similar to each other while clusters formed from these observations are dissimilar from each other. Clustering is mostly used for data mining and/or further analysis for future inferences.

Clustering algorithms are data-type specific. In this case the data has both numerical and categorical observations hence the appropriate algorithm for use is K-prototypes. K-prototypes is an extension of both K-means and K-modes algorithm. 

K-prototypes belongs to  the family of partitional cluster algorithms. It measures distance between numerical features using Euclidean distance (like K-means) and distance between categorical features using the number of matching categories (like K-modes). It was first published by Huang in 1998[^2]. 

Prior to clustering, one needs to derive optimal number of clusters. Since there's no reproducible research on this using K-prototypes, the code below iterates 
probable number of clusters by plotting a scree plot using the elbow method[^3]. 

```{r elbow plot, echo=FALSE, fig.cap="fig 2.0 Scree plot showing the optimum number of clusters", fig.margin=T, message=FALSE, warning=FALSE, paged.print=FALSE}
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```
From figure 2.0, I'll cluster my data using 3 partitions.

```{r k-prototype implementation, eval=FALSE, include=FALSE}
ict_clusters <- kproto(ICTSkills, k=4, lambda = NULL, iter.max = 100, nstart = 1, na.rm = TRUE, verbose = T)
ict_clusters
```

## Visualisation of the clusters

```{r cluster vis, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
clprofiles(ict_clusters, ICTSkills)
```

## Knowledge discovery
Clusters derived include: 

* Gender is constant across the clusters and its independent of the ICT skill. 
* UNICEF regions; Eastern and Southern Africa (ESA),  South Asia (SA) and West and Central Africa (WCA) that have fall under wealth quantile of 1 have the highest count of youths and adults with ICT skills.
* UNICEF regions; East Asia and the Pacific (EAP) had more males who had knowledge in ICT skills and fall under wealth quantile 2, 3 and 4.

## Perfomance comparison

A performance measure is needed to be performed in the underlying clusters because the previous output has clustered only categorical variables using K-modes and numeric variables using k-means. This comparison called Rand Index (Rand measure). 
Rand index is a way of comparing similarity of results between two clustering methods. In this case we want to compare the similarity of results from K-means and K-modes. 


# References

[^1]: https://www.itl.nist.gov/div898/handbook/eda/section3/histogr6.htm
[^2]: https://link.springer.com/article/10.1023/A:1009769707641
[^3]: https://stats.stackexchange.com/questions/293877/optimal-number-of-clusters-using-k-prototypes-method-in-r

